---
title: "WrightMap Tutorial - Part 1"
subtitle: "Plotting Unidimensional/Dichotomous Models"
author: "David Torres Irribarra"
date: "09/25/2024"
toc: true
format: html
css: custom.css
editor: visual
lightbox: true
---

```{r, include=FALSE, message=FALSE}
# Set the CRAN mirror to avoid errors when installing packages
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Silently load the WrightMap package
if (!requireNamespace("WrightMap", quietly = TRUE)) {
  install.packages("WrightMap")

# Silently load the WrightMap package
if (!requireNamespace("RColorBrewer", quietly = TRUE)) {
  install.packages("RColorBrewer")
}

}
library(WrightMap)
```

## Plotting Unidimensional/Dichotomous Models
###### Updated September 2024 for changes in version 1.4.

### Setup and Basics

This is an introduction to the `wrightMap` function in the `WrightMap` package. The wrightMap function creates Wright Maps based on person estimates and item parameters produced by an item response analysis. The CQmodel function reads output files created using ConQuest software and creates a set of data frames for easy data manipulation, bundled in a CQmodel object. The wrightMap function can take a CQmodel object as input or it can be used to create Wright Maps directly from data frames of person and item parameters.

### Setup

Let's start by installing the latest version of the package from CRAN.

```{r}
library(WrightMap)
```

We will also load RColorBrewer to take advantage of its color palettes.

```{r}
library(RColorBrewer)
```

### A Simple Dichotomous Example

To plot a simple Rasch model, we start by creating mock person and item estimates.

For the person proficiencies, we create a single vector with 1000 values.

```{r}
set.seed(2020)
rasch.sim.thetas <- rnorm(1000)
```

And for the item difficulties, we create a vector with 10 values.

```{r}
rasch.sim.thresholds <- runif(10, -3, 3)
```

We now have all we need to create a WrightMap with a single line.

```{r}
#| label: t01-i01-first_example
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds)
```

We can start to customize the Wright Map by simply relabeling its main parts using `main.title`, `axis.logits`, `axis.persons`, and `axis.items`.

```{r}
#| label: t01-i02-first_example_relabel
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds,
    main.title = "This is my example Wright Map",
    axis.persons = "This is the person distribution",
    axis.items = "These are my survey questions")
```

### The Person Side and the Item Side

If you do not like histograms, WrightMap has the option `person.side` that allows you to switch between histograms (the default option: `personHist`) or density using the option `personDens`.

```{r}
#| label: t01-i03-personDens
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, person.side = personDens)
```

Or, you may want to change the way items are represented by using the option `item.side`, which offers `itemModern` (the default representation), `itemClassic` (for ConQuest-style Wright Maps), and a `itemHist` for a histogram summary of the items.

The `itemClassic` option is well suited for cases where you want to include many items using less space.

```{r}
#| label: t01-i04-itemClassic
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
rasch.sim.thresholds.2 <- runif(150, -3, 3)
wrightMap(rasch.sim.thetas, rasch.sim.thresholds.2, item.side = itemClassic)
```

Alternatively, you can represent use a back-to-back histogram representation with `itemHist` (notice that in the following example we are using the option `item.prop` to adjust the relative sizes of the person and item side).

```{r}
#| label: t01-i05-itemHist
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
rasch.sim.thresholds.3 <- rnorm(150)
wrightMap(rasch.sim.thetas, rasch.sim.thresholds.2, item.side = itemHist, item.prop = 0.5)
```

### Customizing Wright Maps

Let us focus on the default item and person side representations to explore their potential customizations.

You might want to remove the label for the item difficulties by setting `show.thr.lab = FALSE`.

```{r}
#| label: t01-i06-nolabels
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.lab = FALSE)
```

Or you might want to see just labels, by turning off symbols with `show.thr.sym = FALSE`.

```{r}
#| label: t01-i07-justlabels
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.sym = FALSE)
```

### Customizing Symbols

Let's start by making all the symbols bigger with `thr.sym.cex = 2.5` (default is 1).

```{r}
#| label: t01-i08-editsymbs
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds,
    show.thr.lab = FALSE, thr.sym.cex = 2.5)
```

To select what kind of symbols you want to use, you can use the `thr.sym.pch` parameter.

```{r}
#| label: t01-i09-editsymbsPCH
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.lab = FALSE,
    thr.sym.cex = 2.5, thr.sym.pch = 17)
```

### Customizing Colors and Labels

We will need some colors for the next examples.

```{r}
#| label: t01-i10-colors
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
display.brewer.pal(10, "Paired")
itemcolors <- brewer.pal(10, "Paired")
```

Now, let's use those colors in our item difficulty symbols using the `thr.sym.col.fg`.

```{r}
#| label: t01-i11-symbcolors
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.lab = FALSE,
    thr.sym.pch = 17, thr.sym.cex = 2.5, thr.sym.col.fg = itemcolors)
```

### Customizing Labels

Define some item names:

```{r}
itemnames <- c("Dasher", "Dancer", "Prancer", "Vixen", "Comet", "Cupid", "Donner", 
    "Blitzen", "Rudolph", "Olive")
```

Now assign them to the item difficulties:

```{r}
#| label: t01-i13-custlabels
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.sym = FALSE,
    thr.lab.text = itemnames)
```

You can also control the size of the labels using the `thr.lab.cex` parameter.

```{r}
#| label: t01-i14-labelsizes
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.sym = FALSE,
    thr.lab.text = itemnames, thr.lab.cex = 1.5)
```

And we can of course control the colors using `thr.lab.col`.

```{r}
#| label: t01-i15-labelcols
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.sym = FALSE,
    thr.lab.text = itemnames, thr.lab.col = itemcolors, thr.lab.cex = 1.5)
```

Finally, if you want to go crazy, you can also change the type style using `thr.lab.font`. This parameter follows the R convention where 1 = plain, 2 = bold, 3 = italic, 4 = bold italic.

```{r}
#| label: t01-i16-labelfont
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch.sim.thresholds, show.thr.sym = FALSE,
    thr.lab.text = itemnames, thr.lab.col = itemcolors,
    thr.lab.cex = 1.5, thr.lab.font = c(1:4))
```