---
title: "WrightMap Tutorial - Part 3"
subtitle: "Plotting Multidimensional & Polytomous Models"
author: "David Torres Irribarra"
date: "25/09/2024"
toc: true
format: html
editor: visual
---

```{r, include=FALSE, message=FALSE}
# Set the CRAN mirror to avoid errors when installing packages
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Silently load the WrightMap package
if (!requireNamespace("WrightMap", quietly = TRUE)) {
  install.packages("WrightMap")
}
library(WrightMap)
```

## Plotting Multidimensional & Polytomous Models
###### Updated September 2024 for changes in version 1.4.

### Intro

In Part 1, we reviewed how to install the package from GitHub and how to customize unidimensional and dichotomous models. Now in Part 2, we’ll look at graphing some more complicated models.

(See Part 3 for ConQuest integration and making thresholds out of deltas.)

In Part 2, we'll look at graphing some more complicated models. First, let's generate some thresholds for a multidimensional model. This will be a matrix containing five columns of person estimates.

### Multidimensional models

We will need again to load RColorBrewer for this example.

```{r}
install.packages("RColorBrewer")
library(RColorBrewer)
```

```{r}
set.seed(2020)
mdim.sim.thetas <- matrix(rnorm(5000), ncol = 5)
```

Since this will be a dichotomous model, we'll generate a single column for thresholds.

```{r}
mdim.sim.thresholds <- runif(10, -3, 3)
```

Okay, let's see what the Wright Map looks like for this.

```{r}
#| label: t02-i01-first_mdim_example
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(mdim.sim.thetas, mdim.sim.thresholds)
```

That doesn't look right. Let's adjust the proportion of the map's parts.

```{r}
#| label: t02-i02-itemprop
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(mdim.sim.thetas, mdim.sim.thresholds, item.prop = 0.5)
```

Let's change the dimensions names.

```{r}
#| label: t02-i03-dimnames
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(mdim.sim.thetas, mdim.sim.thresholds, item.prop = 0.5,
    dim.names = c("Algebra", "Calculus", "Trig", "Stats", "Arithmetic"))
```

And let's give them some color.

```{r}
#| label: t02-i04-dimcolors
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(mdim.sim.thetas, mdim.sim.thresholds, item.prop = 0.5,
    dim.names = c("Algebra", "Calculus", "Trig", "Stats", "Arithmetic"),
    dim.color = brewer.pal(5, "Set1"))
```

And let's associate the items with each dimension.

```{r}
#| label: t02-i05-dimitems
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(mdim.sim.thetas, mdim.sim.thresholds, item.prop = 0.5,
    dim.names = c("Algebra", "Calculus", "Trig", "Stats", "Arithmetic"),
    dim.color = brewer.pal(5, "Set1"), show.thr.lab = FALSE,
    thr.sym.col.fg = rep(brewer.pal(5, "Set1"), each = 2),
    thr.sym.col.bg = rep(brewer.pal(5, "Set1"), each = 2),
    thr.sym.cex = 2, use.hist = FALSE)
```

### Polytomous models

All right, let's look at a Rating Scale Model. First, let's generate three dimensions of person estimates.

```{r}
rsm.sim.thetas <- data.frame(d1 = rnorm(1000, mean = -0.5, sd = 1), d2 = rnorm(1000, 
    mean = 0, sd = 1), d3 = rnorm(1000, mean = +0.5, sd = 1))
```

Now let's generate the thresholds for the polytomous items. We'll make them a matrix where each row is an item and each column a level.

```{r}
items.loc <- sort(rnorm(10))

rsm.sim.thresholds <- data.frame(l1 = items.loc - 1, l2 = items.loc - 0.5,
    l3 = items.loc + 0.5, l4 = items.loc + 1)

rsm.sim.thresholds
```

Let's look at the Wright Map!

```{r}
#| label: t02-i06-polymexample
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rsm.sim.thetas, rsm.sim.thresholds)
```

Let's assign a color for each level.

```{r}
itemlevelcolors <- matrix(rep(brewer.pal(4, "Set1"), 10), byrow = TRUE, ncol = 4)

itemlevelcolors
```

And now make a Wright Map with them.

```{r}
#| label: t02-i07-polymcolors
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rsm.sim.thetas, rsm.sim.thresholds, thr.sym.col.fg = itemlevelcolors,
    thr.sym.col.bg = itemlevelcolors)
```

But we also want to indicate which dimension they belong to... with symbols.

```{r}
itemdimsymbols <- matrix(c(rep(16, 12), rep(17, 12), rep(18, 16)),
    byrow = TRUE, ncol = 4)

itemdimsymbols
```

```{r}
#| label: t02-i08-polymexamplesymb
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rsm.sim.thetas, rsm.sim.thresholds, show.thr.lab = FALSE,
    thr.sym.col.fg = itemlevelcolors, thr.sym.col.bg = itemlevelcolors,
    thr.sym.pch = itemdimsymbols, thr.sym.cex = 2)
```

Additionally, we may want to clearly indicate which item parameters are associated with each item. We can draw lines that connect all parameters connected to an item using the `vertLines` parameter.

```{r}
#| label: t02-i09-vertLinesexample
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rsm.sim.thetas, rsm.sim.thresholds, show.thr.lab = FALSE,
    thr.sym.col.fg = itemlevelcolors, thr.sym.col.bg = itemlevelcolors,
    thr.sym.pch = itemdimsymbols, thr.sym.cex = 2, vertLines = TRUE)
```

### Changing the item axis

What happens if you have too many items?

```{r}
rasch2.sim.thresholds <- runif(50, -3, 3)
```

We use the defaults...

```{r}
#| label: t02-i09-manyitemsdef
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rnorm(1000), rasch2.sim.thresholds)
```

Some options...

You can use the `itemClassic` or `itemHist` options for `item.side`.


```{r}

rasch.sim.thetas <- rnorm(1000)
rasch2.sim.thresholds <- runif(10, -3, 3)

```

```{r}
#| label: t02-i10-manyitemsclassic
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 6
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch2.sim.thresholds, item.side = itemClassic,
    item.prop = .5)
```

```{r}
#| label: t02-i11-manyitemhist
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rasch.sim.thetas, rasch2.sim.thresholds, item.side = itemHist, item.prop = 0.5)
```

Or you can play with the way labels are presented:

```{r}
#| label: t02-i12-manyitem45
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rnorm(1000), rasch2.sim.thresholds, show.thr.lab = FALSE, label.items.srt = 45)
```

```{r}
#| label: t02-i13-manyitem2rows
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rnorm(1000), rasch2.sim.thresholds, show.thr.lab = FALSE, label.items.rows = 2)
```

```{r}
#| label: t02-i14-manyitem3rows
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rnorm(1000), rasch2.sim.thresholds, show.thr.lab = FALSE,
    label.items = c(1:50), label.items.rows = 3)
```

Or you can get rid of that axis completely.

```{r}
#| label: t02-i15-manyitemnoaxis
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(rnorm(1000), rasch2.sim.thresholds, show.thr.sym = FALSE,
    thr.lab.text = paste("I", 1:50, sep = ""), label.items = "",
    label.items.ticks = FALSE)
```