---
title: "WrightMap Tutorial - Part 4"
subtitle: "Using Conquest Output and Making Thresholds"
author: "David Torres Irribarra"
date: "25/09/2024"
toc: true
format: html
editor: visual
---

```{r, include=FALSE, message=FALSE}

# Silently load the WrightMap package
if (!requireNamespace("WrightMap", quietly = TRUE)) {
  install.packages("WrightMap")
}
library(WrightMap)
```

## Using Conquest Output and Making Thresholds
###### Updated September 2024 for changes in version 1.4.

### Intro

In this part of the tutorial, we'll show how to load ConQuest output to make a CQmodel object and then WrightMaps. We'll also show how to turn deltas into thresholds. All the example files here are available in the `/inst/extdata` folder of our GitHub site. If you download the latest version of the package, they should be in a folder called `/extdata` wherever your R packages are stored. You can set this folder as your working directory with `setwd()` or use the `system.file()` command---as in the next set of examples---to run them.

### Making the model

Let's load a model. The first parameter should be the name of the person estimates file, while the second should be the name of the show file. Both are necessary for creating Wright maps (although the `CQmodel` function will run fine with only one or the other, provided that they are properly passed). 

We start by defining a path to the WrightMap example files.

```{r}
fpath <- system.file("extdata", package="WrightMap")
```

And we load the example output.
