---
title: "WrightMap Tutorial - Part 4"
subtitle: "Using Conquest Output and Making Thresholds"
author: "David Torres Irribarra"
date: "09/25/2024"
toc: true
format: html
editor: visual
lightbox: true
---

```{r, include=FALSE, message=FALSE}

# Silently load the WrightMap package
if (!requireNamespace("WrightMap", quietly = TRUE)) {
  install.packages("WrightMap")
}
if (!requireNamespace("this.path", quietly = TRUE)) {
  install.packages("this.path")
}
library(WrightMap)
```


## Using Conquest Output and Making Thresholds
###### Updated March 18, 2016 for changes in version 1.2.

### Intro

In this part of the tutorial, we'll show how to load ConQuest output to make a CQmodel object and then WrightMaps. We'll also show how to turn deltas into thresholds. All the example files here are available in the `/inst/extdata` folder of our GitHub site. If you download the latest version of the package, they should be in a folder called `/extdata` wherever your R packages are stored. You can set this folder as your working directory with `setwd()` or use the `system.file()` command---as in the next set of examples---to run them.

### Making the model

Let's load a model. The first parameter should be the name of the person estimates file, while the second should be the name of the show file. Both are necessary for creating Wright maps (although the `CQmodel` function will run fine with only one or the other, provided that they are properly passed). 

We start by loading the WrightMap example files.

```{r}
fpath <- system.file("extdata", "ex2.SHW",package="WrightMap")
```

And we load the example output.

```r
model1 <- CQmodel(p.est = (system.file("extdata", "ex2.eap",package="WrightMap"))
    , show = (system.file("extdata", "ex2.SHW",package="WrightMap")))
```

```{r}
#| echo: false
#| output: false

model1 <- CQmodel(p.est = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2.eap")              
       , show = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2.shw")
       )

```

This (`model1`) is a `CQmodel` object. Enter the name of the object to see the names of all the tables & information stored within this object.

```{r}
model1
```

Type the name of any of these tables to see the information stored there.

```{r}
model1$SOE
model1$equation
model1$reg.coef
model1$rel.coef
model1$variances
```

The most relevant for our purposes are the `RMP`, `GIN`, and `p.est` tables. The `RMP` tables contain the Response Model Parameters. These are item parameters. Typing `model1$RMP` would display them, but they're a little long, so I'm just going to ask for the names and then show the first few rows of each table.

```{r}
names(model1$RMP)
```

For this model, the RMPs have item and item*step parameters. We could add these to get the deltas. Let's see what the tables look like.

```{r}
head(model1$RMP$item)
head(model1$RMP$"item*step")
```

Let's look at a more complicated example.

```{r}
#| echo: false
#| output: false

model2 <- CQmodel(p.est = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex4a.mle")              
       , show = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex4a.shw")
       )

```

```r
model2 <- CQmodel(file.path(fpath, "ex4a.mle"), file.path(fpath, "ex4a.shw"))
```

```{r}
model2$equation
names(model2$RMP)
head(model2$RMP$"rater*topic*criteria*step")
```

The GIN tables show the threshold parameters.

```{r}
model1$GIN
model2$GIN
```

Finally, the `p.est` table shows person parameters.

```{r}
head(model1$p.est)  ##EAPs
head(model2$p.est)  ##MLEs
```

### CQmodel, meet wrightMap

Ok, we have person parameters and item parameters: Let's make a Wright Map.

```{r}
#| label: t03-i01-CQexample00
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model1)
```

The above uses the GIN table as thresholds. But you may want to use RMP tables. For example, if you have an item table and an item*step table, you might want to combine them to make deltas. You could do this yourself, but you could also let the `make.deltas` function do it for you. This function reshapes the item*step parameters, checks the item numbers to see if there are any dichotomous items, and then adds the steps and items. This can be especially useful if you didn't get a GIN table from ConQuest (see below).

```{r}
#| echo: false
#| output: false

model3 <- CQmodel(p.est = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2a.eap")              
       , show = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2a.shw")
       )

```

```r
model3 <- CQmodel(file.path(fpath, "ex2a.eap"), file.path(fpath, "ex2a.shw"))
```

```{r}
model3$GIN
model3$equation
make.deltas(model3)
```

When sent a model with no GIN table, `wrightMap` will automatically send it to `make.deltas` without the user having to ask.

```{r}
#| label: t03-i02-CQexample01
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model3, label.items.row = 2)
```

The `make.deltas` function can also handle rating scale models.

```{r}
#| echo: false
#| output: false

model4 <- CQmodel(p.est = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2b.eap")              
       , show = file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2b-2.shw")
       )

```

```r
model4 <- CQmodel(file.path(fpath, "ex2b.eap"), file.path(fpath, "ex2b-2.shw"))
```

```{r}
model4$GIN
model4$equation
make.deltas(model4)
```

Or let `wrightMap` make them automatically.

```{r}
#| label: t03-i03-CQexample04
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model4, label.items.row = 2)
```

### Specifying the tables

In the above examples, we let `wrightMap` decide what parameters to graph. `wrightMap` starts by looking for a GIN table. If it finds that, it assumes they are thresholds and graphs them accordingly. If there is no GIN table, it then sends the function to `make.deltas`, which will examine the model equation to see if it knows how to handle it. `Make.deltas` can handle equations of the form:

```r
A (e.g. `item`)
A + B (e.g. `item + step` [RSM])
A + A * B (e.g. `item + item * step` [PCM])
A + A * B + B (e.g `item + item * gender + gender`)
```

But sometimes we may want something other than the default. Let's look at `model2` again.

```{r}
model2$equation
```

Here's the default Wright Map, using the GIN table:

```{r}
#| label: t03-i04-CQexample05
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, min.logit.pad = -29)
```

This doesn't look great. Instead of showing all these estimates, we can specify a specific RMP table to use using the `item.table` parameter.

```{r}
#| label: t03-i05-CQexample06
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "rater")
```

That shows just the rater parameters. Here's just the topics.

```{r}
#| label: t03-i06-CQexample07
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "topic")
```

What I really want, though, is to show the rater*topic estimates. For this, we can use the `interactions` and `step.table` parameters.


```{r}
#| label: t03-i07-CQexample08
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "rater", interactions = "rater*topic",
    step.table = "topic")
```

Switch the item and step names to graph it the other way:

```{r}
#| label: t03-i08-CQexample09
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "topic", interactions = "rater*topic",
    step.table = "rater")
```

You can leave out the interactions to have more of a rating scale-type model.

```{r}
#| label: t03-i09-CQexample10
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "rater", step.table = "topic")
```

Or leave out the step table:

```{r}
#| label: t03-i10-CQexample11
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "rater", interactions = "rater*topic")
```

Again, `make.deltas` is reading the model equation to decide whether to add or subtract. If, for some reason, you want to specify a different sign for one of the tables, you can use `item.sign`, `step.sign`, and `inter.sign` for that.

```{r}
#| label: t03-i11-CQexample12
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model2, item.table = "rater", interactions = "rater*topic",
    step.table = "topic", step.sign = -1)
```

### Making thresholds

So far, we've seen how to use the GIN table to graph thresholds, or the RMP tables to graph deltas. We have one use case left: Making thresholds out of those RMP-generated deltas. The `make.thresholds` function can handle this. The example below uses the `model3` deltas, but you can send it any matrix with items as rows and steps as columns.

```{r}
deltas <- make.deltas(model3)
deltas
make.thresholds(deltas)
```

Alternately, we can just send the model object directly:

```{r}
make.thresholds(model3)
```

You don't have to do any of this to make a Wright Map. You can just send the model to `wrightMap`, and use the `type` parameter to ask it to calculate the thresholds for you.

```{r}
#| label: t03-i12-CQexample13
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(model3, type = "thresholds", label.items.row = 2)
```

Finally: If all you want is the Wright Maps, you can skip `CQmodel` entirely and just send your files to `wrightMap`.

```r
wrightMap(file.path(fpath,"ex2a.eap"), file.path(fpath,"ex2.shw"), label.items.row = 3)
```

```{r}
#| label: t03-i13-CQexample14
#| dev: ["svg", "png"]
#| fig-cap: ""
#| results: "hide"
#| fig-width: 8
#| fig-height: 5
#| dpi: 300
wrightMap(
    file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2a.eap"), 
    file.path("https://raw.githubusercontent.com/david-ti/wrightmap/refs/heads/master/inst/extdata/ex2a.shw")
        , label.items.row = 3)
```